---
title: "SAFE examples - apartments dataset"
author: "Anna Gierlak"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(SAFE)
library(DALEX)
# library(changepoint.np)
# library(strucchange)
library(randomForest)
# library(ggplot2)
library(gridExtra)
library(dplyr)
# library(utils)
# library(stringr)
# library(iml)
# library(factorMerger)
```

Fitting models to the original dataset - linear model and random forest:

```{r, warning = FALSE}
model_lm1 <- lm(m2.price ~ ., data = apartments)
explainer_lm1 <- explain(model_lm1, data = apartmentsTest[1:3000,2:6], y = apartmentsTest[1:3000,1], label = "lm1")
set.seed(111)
model_rf1 <- randomForest(m2.price ~ ., data = apartments)
explainer_rf1 <- explain(model_rf1, data = apartmentsTest[1:3000,2:6], y = apartmentsTest[1:3000,1], label = "rf1")
```


Creating safe_extractor object using SAFE package:

```{r, warning = FALSE}
safe_extractor <- safe_extraction(explainer_rf1, verbose = FALSE)
```

```{r, warning = FALSE}
print(safe_extractor)
```

In the plot above we can see which points have been chosen to be the breakpoints for particular variable: 

```{r, fig.width=6, warning = FALSE}
plot(safe_extractor, "construction.year")
```

Using created object to extract new features:

```{r, fig.width=6, warning = FALSE}
data1_categorical <- safely_transform_data(safe_extractor, apartmentsTest[3001:6000,], encoding = "categorical", verbose = FALSE)
```

```{r, echo = FALSE}
knitr::kable(head(data1_categorical))
```

The same with one-hot encoding:

```{r, fig.width=6, warning = FALSE}
data1_onehot <- safely_transform_data(safe_extractor, apartmentsTest[3001:6000,], encoding = "one-hot", verbose = FALSE)
```

```{r, echo = FALSE}
knitr::kable(head(data1_onehot))
```

Feature selection using both "categorical" and "one-hot" methods of encoding:

```{r, fig.width=6, warning = FALSE}
vars_categorical <- safely_select_variables(safe_extractor, data1_categorical, which_y = 2, encoding = "categorical", verbose = FALSE)
print(vars_categorical)
vars_onehot <- safely_select_variables(safe_extractor, data1_onehot, which_y = 2, encoding = "one-hot", verbose = FALSE)
print(vars_onehot)

data1_categorical <- data1_categorical[,c('m2.price', vars_categorical)]
data1_onehot <- data1_onehot[,c('m2.price', vars_onehot)]

```

First few rows for data with categorical encoding, after feature selection:

```{r, echo = FALSE}
knitr::kable(head(data1_categorical))
```

Transforming another data that will be used later in explainers:

```{r, fig.width=6, warning = FALSE}
data2_categorical <- safely_transform_data(safe_extractor, apartmentsTest[6001:9000,], encoding = "categorical", verbose = FALSE)[,c('m2.price', vars_categorical)]
data2_onehot <- safely_transform_data(safe_extractor, apartmentsTest[6001:9000,], encoding = "one-hot", verbose = FALSE)[,c('m2.price', vars_onehot)]
```




Fitting models to data containg newly created columns:

```{r, warning = FALSE}
model_lm2_categorical <- lm(m2.price ~ ., data = data1_categorical)
explainer_lm2_categorical <- explain(model_lm2_categorical, data = data2_categorical, y = apartmentsTest[6001:9000,1], label = "lm2 - categorical")

model_lm2_onehot <- lm(m2.price ~ ., data = data1_onehot)
explainer_lm2_onehot <- explain(model_lm2_onehot, data = data2_onehot, y = apartmentsTest[6001:9000,1], label = "lm2 - onehot")

set.seed(111)
model_rf2_categorical <- randomForest(m2.price ~ ., data = data1_categorical)
explainer_rf2_categorical <- explain(model_rf2_categorical, data2_categorical, apartmentsTest[6001:9000,1], label = "rf2 - categorical")

set.seed(111)
model_rf2_onehot <- randomForest(m2.price ~ ., data = data1_onehot)
explainer_rf2_onehot <- explain(model_rf2_onehot, data2_onehot, apartmentsTest[6001:9000,1], label = "rf2 - onehot")
```


```{r, echo = FALSE, warning = FALSE}
mp_lm1 <- model_performance(explainer_lm1)
mp_rf1 <- model_performance(explainer_rf1)
mp_lm2_categorical <- model_performance(explainer_lm2_categorical)
mp_rf2_categorical <- model_performance(explainer_rf2_categorical)
mp_lm2_onehot <- model_performance(explainer_lm2_onehot)
mp_rf2_onehot <- model_performance(explainer_rf2_onehot)
```

```{r, echo = FALSE, fig.width=7, fig.height=6, warning = FALSE}
plot(mp_lm1, mp_rf1, mp_lm2_categorical, mp_rf2_categorical, mp_lm2_onehot, mp_rf2_onehot, geom = "boxplot")
```


